---
output: github_document
---


# dpukarchive
<!-- badges: start -->
<!-- badges: end -->

Lots of police data is made available for public use at [data.police.uk](https://data.police.uk). The site offers an API - but, on exploration, it seems like the API only gives access to data back as far as 2021. The site also has an ['archive' page](https://data.police.uk/data/archive/) - the archive hosts data from Dec 2010 onwards. <br>

However, the archive is difficult to use. It hosts a number of zipped folders, available for download. Each zipped folder contains a number of subfolders, which contain one month's worth of data. For every month, there should be one csv for each police force, for each data series (crime, outcome, stop & search data), though some are missing. In sum, there are lots of files inside folders inside folders, and to extract the particular data files you want to use presently involves downloading and manually searching through many files you don't want. <br>

Plan: to try develop a package that makes interacting with the UK police data archive easier. 
<!-- I'm trying to take stock of the archive - figure out what data files are absent/present/present in multiple versions, how to interact with the archive in the most efficient way, etc.results='hide',fig.keep='all'-->
 
## Understanding what data is present in the archive
Have checked the contents of each zipped folder on the page, and recorded that information in a table (code that produces the table is in file [archive_contents.R](https://github.com/katehayes/dpukarchive/blob/main/data-raw/archive_contents.R) and the table itself is in the data folder). This info will feed into package functions. 

```{r archive_contents, echo = F, message = F}
library(dplyr)
library(magrittr)

load("data/archive_contents.Rda") 

archive_contents %>% 
  mutate(year_month = paste(year, month, sep="-")) %>% 
  group_by(folder) %>% 
  mutate(folder_min = min(year_month[contains == T]),
         folder_max = max(year_month[contains == T])) %>% 
  ungroup() %>% 
  mutate(should_contain = ifelse(year_month >= folder_min & year_month <= folder_max,
                          T, F)) %>% 
  select(file_name, folder, should_contain, contains) %>% 
  head()

```
 


```{r availability-plot, echo = F, message = F, dev='png', dpi = 1000, dev.args = list(bg = 'transparent')}
library(dplyr)
library(magrittr)
library(ggplot2)

force_list <- httr::GET("https://data.police.uk/api/forces") %>%
  httr::content("text") %>% # extract the JSON
  jsonlite::fromJSON() # convert the JSON string to a list

 delete_no_display <- function(v) {
    if_else(stringr::str_detect(v, 'no_display'), '', v)
  }

load("data/archive_contents.Rda")

# lets do ever, always, mever
archive_contents %>% 
  mutate(year_month = paste(year, month, sep="-")) %>% 
  group_by(folder) %>% 
  mutate(folder_min = min(year_month[contains == T]),
         folder_max = max(year_month[contains == T])) %>% 
  ungroup() %>% 
  mutate(should_contain = ifelse(year_month >= folder_min & year_month <= folder_max,
                          T, F)) %>% 
  filter(should_contain == T) %>% 
  group_by(file_name) %>% 
  mutate(file_present = if_else(all(contains == T), "never missing.", if_else(all(contains == F), "always missing;", "sometimes missing;"))) %>% 
  mutate(file_present = factor(file_present, levels = c("always missing;", "sometimes missing;", "never missing."))) %>% 
  ungroup() %>% 
  distinct(series, police, file_name, file_present) %>% 
  mutate(year_month = substr(file_name, 1, 7),
         series = factor(series, levels = c("street", "outcomes", "stop-and-search"))) %>% 
  left_join(force_list %>% 
              rename(police = id)) %>% 
  mutate(police_label = ifelse(police == "metropolitan" & series == "stop-and-search", name, paste0(name, 'no_display')),
         police_label = ifelse(police == "avon-and-somerset" & series == "stop-and-search", name, police_label),
         police_label = ifelse(police == "cheshire" & series == "stop-and-search", name, police_label),
         police_label = ifelse(police == "derbyshire" & series == "stop-and-search", name, police_label),
         police_label = ifelse(police == "greater-manchester" & series == "stop-and-search", name, police_label),
         police_label = ifelse(police == "west-midlands" & series == "stop-and-search", name, police_label)) %>% 
  # mutate(police_label = ifelse(series == "street", name, paste0(name, 'no_display'))) %>%
  ggplot() +
  geom_tile(aes(x = year_month, y = forcats::fct_rev(police_label), fill = file_present),
            colour = "black") +
  facet_wrap(~series, ncol = 1, scale = 'free_y',
             labeller = as_labeller(c(street = "Street", 
                                      outcomes = "Outcomes", 
                                      `stop-and-search` = "Stop & search")), 
             strip.position = "left"
             ) +
  # scale_fill_manual(values = c("#20c45b", "#fc4724", "#fc7d24"))+
  scale_fill_manual(values = c("#adabaa", "#fc7d24", "#fc4724"))+
  scale_x_discrete(name = "",
                   breaks = c("2011-01", "2012-01", "2013-01",
                              "2014-01", "2015-01", "2016-01", "2017-01", "2018-01",
                              "2019-01", "2020-01", "2021-01", "2022-01",
                              "2023-01", "2024-01"),
                   labels = c("2011", "2012", "2013", 
                              "2014", "2015", "2016", "2017", "2018",
                              "2019", "2020", "2021", "2022",
                              "2023", "2024"),
                   expand = c(0,0)) +
  scale_y_discrete(position = "right",
                   name = "",
                   labels = delete_no_display,
                   breaks = delete_no_display,
                   # labels = "",
                   expand = c(0,0)) +
  theme(strip.background = element_rect(fill='transparent'),
        strip.text = element_text(colour = "white"),
        # strip.placement = "outside",
        plot.background = element_rect(fill='transparent'),
        panel.background = element_rect(fill='transparent'),
        legend.background =  element_rect(fill='transparent'),
        axis.ticks = element_line(colour = "white"),
        # axis.line.x.bottom = element_line(colour = "white"),
        axis.text.y = element_text(size = 7, colour = "white"),
        axis.text.x = element_text(size = 9, colour = "white"),
        legend.position = c(0.64, -0.225),
        legend.title = element_blank(),
        legend.key.size = unit(0.5, "lines"),
        legend.text = element_text(size = 11,
                                   colour = "white"),
        # panel.background = element_rect(fill = NA, color = "black"),
        plot.margin = unit(c(1,1,1,1), "lines"),
        # plot.caption = element_text(hjust=1.137,
        #                             vjust = 100,
        #                             size=rel(1)),
        plot.caption = element_text(hjust = 0,
                                    vjust = 0,
                                    size = 11,
                                    colour = "white"),
        panel.spacing = unit(0.1,'lines')) +   #only v small spaces between facets
  guides(fill=guide_legend(nrow = 1)) +
# coord_cartesian(clip="off") +
labs(caption = "Each tile represents one month of data from one (of three) data series and one police force. \nThe colour of the tile indictates whether the data is present or absent in the folders it\nshould be in:") 

# plot_missing_byfile

```

For more details use function fs_report - specify police force, data series, and time span, and the function returns a plot summarising the presence/absence of the relevant files across the relevant folders. Here's an example of the plot for Wiltshire's stop and search data for the months of Jan 2019 to June 2024.

```{r fs-plot, echo = F, message = F, warning = F, dev='png', dpi = 1000, dev.args = list(bg = 'transparent')}

devtools::install_github("katehayes/dpukarchive", force = T, quiet = T)
library(dpukarchive)

plt <- dpukarchive:::fs_graph(police_force = "wiltshire", data_series = "stop-and-search", month_min = "2019-01", month_max = "2024-06")

plt
```

## Extracting data from the police data archive

Function arc_extract -- should be able to specify police force, data series, and time span, and have a dataframe returned efficiently. 

```{r example}
## library(dpukarchive)
## basic example code
```
 
 
## Installation

You can install the development version of dpukarchive from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
# devtools::install_github("katehayes/dpukarchive")
```





